---
title: "Predictive skill"
author: "Timo Kelder"
date: "October 19, 2019"
output: github_document 
---

```{r eval=FALSE, include=FALSE}
Rscript -e "rmarkdown::render('Predictive_skill.Rmd')"
```

In this notebook, we show the forecasts of the Norwegian West Coast compared to observed values. 

## Import data and packages

```{r}
# dir='//home/timok/timok/SALIENSEAS/SEAS5/ensex'
# plotdir=paste0(dir,'/statistics/multiday/plots')
# dir='/home/timok/ensex'
# plotdir='/home/timok/Documents/ensex/R/graphs'
dir='C:/Users/gytk3/OneDrive - Loughborough University/GitHub/EnsEx/Data'
source('Load_data.R')
```

##SEAS5 compared to SeNorge

Plot the observations and the forecasts for: 

- The raw values for the observations and the forecasts
- The anomalies
- The standardized anomalies 

First, we define some functions
```{r fig1, fig.height = 9, fig.width = 5, fig.align = "center"}

rank.histogram <- function(pred,obs=NULL) {
  
  N <- dim(pred)[2]
  K <- dim(pred)[1]
  
  ranks <- apply(rbind(obs, pred), 2, rank, ties.method="random")[1, ]
  rank.hist <- hist(ranks, breaks=seq(-2.5, K+2.5,5),main=NULL)[["counts"]]
}

pred=apply(Extremes_WC,MARGIN = c(1,2) , FUN=calc_anomaly)

rank.histogram3 <- function(pred,obs=NULL) {
  N <- dim(pred)[1]
  K <- dim(pred)[2]
  
  ranks <- apply(cbind(obs, pred), 1, rank, ties.method="random")[1, ]
  rank.hist <- hist(ranks, breaks=seq(-2.5, K+2.5,5),main=NULL)[["counts"]]
}

```

To compare the obervation with the simulations, we show the rank histogram for each of the three plots. The rank histograms illustrates on what rank the observations lie compared to all forecasts. If the simulations are not biased and represent plausible realizations of reality, the observations would be randomly distributed over the range of simulations and the rank histogram would be flat. 

From both the time series and histogram plots, it is clear that the forecast have a lower bias. When corrected for this mean bias, the ranks of the observations seem to be equally distributed over the forecast ranks.   

```{r}
##Plot
# png('../graphs/Rank_hist.png', width = 6, height = 6, units = 'in', res = 300)
par(mfrow=c(3,2),mar=c(1,4,2,0))
plot(1981:2015,predictand,type='l',ylim=c(0,175), ylab = 'SON-3DP events (mm)',xlab = '')
for (mbr in 1:25){
  for (ld in 1:4){
    lines(1981:2015,Extremes_WC[mbr,ld,],col=alpha('blue',0.1))}}
axis(side = 1,1981:2015,labels = F)

par(mar=c(1,6,2,1))

rank1=rank.histogram(rbind(Extremes_WC[,1,],Extremes_WC[,2,],Extremes_WC[,3,],Extremes_WC[,4,]),predictand)

par(mar=c(1,4,2,0))

plot(1981:2015,predictand-mean(predictand),type='l',ylim=c(-60,80),ylab = 'Anomalies (mm)',xlab = '')
for (mbr in 1:25){
  for (ld in 1:4){
    lines(1981:2015,Extremes_WC[mbr,ld,]-mean(Extremes_WC),col=alpha('blue',0.1))}}
axis(side = 1,1981:2015,labels = F)


par(mar=c(1,6,2,1))

rank2=rank.histogram(rbind(Extremes_WC[,1,],Extremes_WC[,2,],Extremes_WC[,3,],Extremes_WC[,4,])-mean(Extremes_WC),predictand-mean(predictand))

par(mar=c(2,4,2,0))

plot(1981:2015,predictand_anomaly,type='l',ylim = c(-2.5,4),ylab='Standardized anomalies (-)',xlab = '')
for (mbr in 1:25){
  for (ld in 1:4){
    lines(1981:2015,calc_anomaly(Extremes_WC[mbr,ld,]),col=alpha('blue',0.1))}}
axis(side = 1,1981:2015,labels = F)

par(mar=c(2,6,2,1))

rank3=rank.histogram3(cbind(pred[,,1],pred[,,2],pred[,,3],pred[,,4]),predictand_anomaly)

# dev.off()
```


